---
title: "Group-Ass1- technical report"
author: "Kunal, Jacob and Devanshi"
date: "2024-08-26"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Installing required libraries and preparing the environment

```{r}

options(repos = "https://cran.rstudio.com/")
install.packages("tidyverse")
install.packages("dyplr")
install.packages("ggplot2")
install.packages("mice")
install.packages("knitr")
install.packages("corrplot")
install.packages("reshape2")
install.packages("janitor")
install.packages("imputeTS")
install.packages("skimr")
library(skimr)
library(imputeTS)
library(janitor)
library(tidyverse)
library(ggplot2)
library(mice)
library(knitr)
library(corrplot)
library(reshape2)
library(forecast)
library(tseries)


```

```{r}
#set working directory

setwd("~/Documents/Data Intelligence")

```

#### Importing the datasets

```{r}

gdp <- read_csv("Total-GDP-NZ.csv", show_col_types = FALSE)
total_co2 <- read_csv("Total-CO2.csv", show_col_types = FALSE)
ann_co2 <- read_csv("Ann-CO2.csv", show_col_types = FALSE)
ann_ghg <- read_csv("Ann-GHG-NZ.csv", show_col_types = FALSE)
co2_sector <- read_csv("CO2-sector.csv", show_col_types = FALSE)
gdp_pc <- read_csv("GDP-PC-NZ.csv", show_col_types = FALSE)
ghg_gas <- read_csv("GHG-Gas.csv", show_col_types = FALSE)
ghg_sector <- read_csv("GHG-sector.csv", show_col_types = FALSE)
ghg_pc <- read_csv("GHG_per_capita.csv", show_col_types = FALSE)

```

#### Descriptive Statistics of all the datasets

```{r}
#Annual GDP

head(gdp)
tail(gdp)
summary(gdp)

```

```{r}
#Cumulative annual CO2

head(total_co2)
tail(total_co2)
summary(total_co2)

```

```{r}

#Annual CO2

head(ann_co2)
tail(ann_co2)
summary(ann_co2)
```

```{r}
#Annual GHG

head(ann_ghg)
tail(ann_ghg)
summary(ann_ghg)

```

```{r}

#Cumulative CO2 emissions from multiple sources NZ

head(co2_sector)
tail(co2_sector)
summary(co2_sector)
```

```{r}

#GDP per capital NZ

head(gdp_pc)
tail(gdp_pc)
summary(gdp_pc$`GDP per capita`)

```

```{r}

#GHG emissions from various gases 

head(ghg_gas)
tail(ghg_gas)
summary(ghg_gas)
```

```{r}

#GHG emissions from various sectors in NZ 
head(ghg_sector)
tail(ghg_sector)
summary(ghg_sector)
```

```{r}
#GHG per capita

ghg_pc <- ghg_pc %>% 
  filter(Entity == "New Zealand")
  
head(ghg_pc)
tail(ghg_pc)
summary(ghg_pc$`Per-capita greenhouse gas emissions in CO₂ equivalents`)


```

#### Data Preparation for Analysis before joining

#Since most of the data is available from 1990 to 2022, we filtered all the datasets to show from 1990 to recent.

```{r}

ann_co2 <- ann_co2 %>%
  filter(ann_co2$Year >= 1990)

```

```{r}
ann_ghg <- ann_ghg %>% 
  filter(Year >= 1990)
```

```{r}
ghg_gas <- ghg_gas %>%
  filter(Year >= 1990)
```

```{r}
total_co2 <- total_co2 %>% 
  filter(Year >= 1990)
```

#### Checking for any missing values in all the datasets

```{r}
sum(is.na(ann_co2))
sum(is.na(ann_ghg))
sum(is.na(co2_sector))
sum(is.na(gdp))
sum(is.na(gdp_pc))
sum(is.na(ghg_gas))
sum(is.na(ghg_sector))
sum(is.na(total_co2))
```

#### Merging the Annual GHG and Sector Wise GHG dataset for predictive analytics techniques like correlation analysis and regression analysis

```{r}

#Preparing the data
ghg1 <- left_join(ghg_gas, ann_ghg, by = "Year") #combining nitroux oxide, methane and CO2 emissions to annula ghg datdset
ghg1 <- clean_names(ghg1)

ghg1$Annual_GHG <- ghg1$annual_greenhouse_gas_emissions_in_co2_equivalents
ghg1$annual_greenhouse_gas_emissions_in_co2_equivalents <- NULL

ghg1 <- ghg1 %>%
  select(year, Annual_GHG, everything()) #rearranging columns by year, Annual GHG and then the other gases

colnames(ghg1) <- c("Year", "Annual_GHG", "Nitrous_Oxide", "Methane", "CO2")

#custom summary table
summary_table <- data.frame(
  Variable = colnames(ghg1)[-1],  # Exclude 'Year'
  N = nrow(ghg1),
  Mean = sapply(ghg1[-1], mean, na.rm = TRUE),
  Median = sapply(ghg1[-1], median, na.rm = TRUE),
  SD = sapply(ghg1[-1], sd, na.rm = TRUE),
  Min = sapply(ghg1[-1], min, na.rm = TRUE),
  Max = sapply(ghg1[-1], max, na.rm = TRUE)
)

# Print the summary
print(summary_table)

```

#### Correlation Analysis to check which gas highly correlates with Annual GHG emissions in NZ

```{r}
colnames(ghg1) #checking column names to make sure all the columns are aligned properly
```

```{r}

#correlation matrix to check which gases have high correlation with Annual GHG emissions

ghg_corr <- ghg1 %>%
  select(Annual_GHG, Nitrous_Oxide, 
         Methane, CO2)

correlation_matrix <- cor(ghg_corr, use = "complete.obs")

kable(correlation_matrix, digits = 2, caption = "Correlation Matrix for GHG Emissions Data") #using kable to represent data in a proper tabular format


```

Correlation heatmap

```{r}

#correlation heatmap of Annual GHG emissions and other gases

correlation_data <- melt(correlation_matrix)

ggplot(data = correlation_data, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "red", high = "green", mid = "yellow", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name = "Correlation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 10, hjust = 1),
        axis.text.y = element_text(size = 12)) +
  labs(title = "Correlation Heatmap for GHG Emissions Data", x = "", y = "")
```

Regression Analysis

```{r}
#regression analysis to check how other gases influence Annual GHG in NZ

colnames(ghg1)
ghg_reg <- lm(Annual_GHG ~  + Annual_GHG + Nitrous_Oxide + Methane + CO2, data = ghg1)
summary(ghg_reg)

kable(summary(ghg_reg)$coefficients, caption = "Regression Coefficients")

#extracting values adjusted R, multiple R and f-statistics from the model 
adjusted_r_squared <- summary(ghg_reg)$adj.r.squared
multiple_r <- summary(ghg_reg)$r.squared
f_statistic <- summary(ghg_reg)$fstatistic[1]

cat("Adjusted R-squared:", adjusted_r_squared, "\n")
cat("Multiple R:", multiple_r, "\n")
cat("F-statistic:", f_statistic, "\n")


#plotting regression 

ghg1$predicted_Annual_GHG <- predict(ghg_reg, newdata = ghg1)

ggplot(ghg1, aes(x = Annual_GHG, y = predicted_Annual_GHG)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  labs(title = "Actual vs. Predicted Annual GHG",
       x = "Actual Annual GHG",
       y = "Predicted Annual GHG")

```

#### Analyzing GDP, Annual GHG, and Annual CO2 emissions to see which variable has a high correlation with GDP in New Zealand

```{r}

#merging gdp, ghg per capita, annual ghg and annual CO2 datasets

gdp1 <- left_join(gdp_pc, ann_ghg, by = "Year") 
gdp1

colnames(gdp1)
gdp1 <- gdp1 %>%
  select(Year, `GDP per capita`, everything())

gdp1 <- left_join(gdp1, ann_co2, by = "Year")
gdp1

#filtering ghg per capita data from the year 1990 to 2022
ghg_pc <- ghg_pc %>%
  filter(Year >= 1990)

gdp1 <- left_join(gdp1, ghg_pc, by = "Year")
gdp1$Entity <- NULL 
gdp1$Code <- NULL
gdp1$predicted_GDP_per_capita <- NULL


colnames(gdp1) <- c("Year", "GDP_per_capita", "Annual_GHG", "Annual_CO2", "GHG_per_capita")

print(gdp1)

```

```{r}
#Descriptive statistics

summary(gdp1)

summary_gdp1 <- data.frame(
  Variable = colnames(gdp1)[-1],  # Exclude 'Year'
  N = nrow(gdp1),
  Mean = sapply(gdp1[-1], mean, na.rm = TRUE),
  Median = sapply(gdp1[-1], median, na.rm = TRUE),
  SD = sapply(gdp1[-1], sd, na.rm = TRUE),
  Min = sapply(gdp1[-1], min, na.rm = TRUE),
  Max = sapply(gdp1[-1], max, na.rm = TRUE)
)

print(summary_gdp1)
```

#### Correlation Analysis of GDP per capita, Annual GHG and CO2 emissions

```{r}
#correlatoion analysis for GDP and Annual GHG Emissions in NZ 

colnames(gdp1)
gdp_corr <- gdp1 %>%
  select(GDP_per_capita,Annual_GHG, Annual_CO2, GHG_per_capita)

gdp_corr_matrix <- cor(gdp_corr, use = "complete.obs")

kable(gdp_corr_matrix, digits = 2, caption = "Correlation Matrix for GDP per capita, GHG, CO2 emissions and GHG per capita Data")


#Correlation heatmap

correlation_data1 <- melt(gdp_corr_matrix)

ggplot(data = correlation_data1, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "red", high = "green", mid = "yellow", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name = "Correlation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 10, hjust = 1),
        axis.text.y = element_text(size = 12)) +
  labs(title = "Correlation Heatmap for GDP per capita, Annual GHG and CO2 Emissions", x = "", y = "")
```

#### regression analysis for gdp, ghg, c02

```{r}

#regression analysis for gdp, ghg, c02 to show how GHG and Annual CO2 impacts GDP per capita in NZ

colnames(gdp1) 

gdp_co_ghg_reg <- lm(GDP_per_capita ~ Annual_GHG + 
                        Annual_CO2 + GHG_per_capita, data = gdp1)
summary(gdp_co_ghg_reg)

kable(summary(gdp_co_ghg_reg)$coefficients, caption = "Regression Coefficients")

#extracting values adjusted R, multiple R and f-statistics from the model 
adjusted_r_squared <- summary(gdp_co_ghg_reg)$adj.r.squared
multiple_r <- summary(gdp_co_ghg_reg)$r.squared

cat("Adjusted R-squared:", adjusted_r_squared, "\n")
cat("Multiple R:", multiple_r, "\n")


```

```{r}
#plotting regression 

gdp1$predicted_GDP_per_capita <- predict(gdp_co_ghg_reg, newdata = gdp1)

# Plot the actual values against the predicted values
ggplot(gdp1, aes(x = `GDP_per_capita`, y = predicted_GDP_per_capita)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  labs(title = "Actual vs. Predicted GDP per capita",
       x = "Actual GDP per capita",
       y = "Predicted GDP per capita")
```

#### Exploring GDP and Sector wise GHG in NZ to see which sector significantly impacts the GDP

```{r}

#merging the GDP and Sector wise GHG emission data

gdp_sec_ghg <- left_join(gdp_pc, ghg_sector, by = "Year") 
gdp_sec_ghg

colnames(ann_ghg) <- c("Year", "Annual_GHG")

gdp_sec_ghg <- left_join(ann_ghg, gdp_sec_ghg, by = "Year")
gdp_sec_ghg

gdp_sec_ghg$Country <- NULL  #removed country column as it is not needed

```

```{r}
#rearranging columns in gdp_sec_ghg

gdp_sec_ghg <- gdp_sec_ghg %>%
  select(Year, `GDP per capita`, everything())
 
```

```{r}

#checking if there are any missing values 

sum(is.na(gdp_sec_ghg))

```

#### correlation analysis between GDP per capita, Annual GHG emissions and GHG emissions from various sectors in NZ

```{r}

#handling missing values

colnames(gdp_sec_ghg)
gdp_sec_ghg <- clean_names(gdp_sec_ghg)

gdp_sec_ghg <- gdp_sec_ghg[1:31, ] #dropped year 2021 and 2022 as no imputation method like mice and amelia worked due to high colinearity between columns
```

```{r}
#descriptive statistics


summary_gdp_sec_ghg <- data.frame(
  Variable = colnames(gdp_sec_ghg)[-1],  # Exclude 'Year'
  N = nrow(gdp_sec_ghg),
  Mean = sapply(gdp_sec_ghg[-1], mean, na.rm = TRUE),
  Median = sapply(gdp_sec_ghg[-1], median, na.rm = TRUE),
  SD = sapply(gdp_sec_ghg[-1], sd, na.rm = TRUE),
  Min = sapply(gdp_sec_ghg[-1], min, na.rm = TRUE),
  Max = sapply(gdp_sec_ghg[-1], max, na.rm = TRUE)
)
print(summary_gdp_sec_ghg)
```

```{r}

#### Correlation analysis #### 

 gdp_sec_ghg_corr <- gdp_sec_ghg %>%
  select(gdp_per_capita, annual_ghg, greenhouse_gas_emissions_from_agriculture, greenhouse_gas_emissions_from_land_use_change_and_forestry, greenhouse_gas_emissions_from_waste, greenhouse_gas_emissions_from_buildings, greenhouse_gas_emissions_from_industry, greenhouse_gas_emissions_from_manufacturing_and_construction, greenhouse_gas_emissions_from_transport, greenhouse_gas_emissions_from_electricity_and_heat, fugitive_emissions_of_greenhouse_gases_from_energy_production, greenhouse_gas_emissions_from_other_fuel_combustion, greenhouse_gas_emissions_from_bunker_fuels)

gdp_sec_ghg_corr_matrix <- cor(gdp_sec_ghg_corr, use = "complete.obs")

kable(gdp_sec_ghg_corr_matrix, digits = 2, caption = "Correlation Matrix for GDP, Annnual GHG, and GHG emissions from multiple sectors in NZ")

```

Correlation heatmap

```{r}

correlation_gdp_ghg_sec <- melt(gdp_sec_ghg_corr_matrix)

ggplot(data = correlation_gdp_ghg_sec, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "red", high = "green", mid = "yellow", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab",
                       name = "Correlation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1, size = 6, hjust = 1),
        axis.text.y = element_text(size = 6)) +
  labs(title = "Correlation Heatmap for GDP, Annual GHG and Sectorwisew GHG Emissions Data", x = "", y = "")

```

Regression analysis for GDP, Annual GHG and Sectorwise emissions in NZ

```{r}

#Regression analysis for GDP, Annual GHG and Sectorwise emissions in NZ 

colnames(gdp_sec_ghg)

gdp_ghg_sec_reg <- lm(gdp_per_capita ~ annual_ghg +
  greenhouse_gas_emissions_from_agriculture +
  greenhouse_gas_emissions_from_land_use_change_and_forestry +
  greenhouse_gas_emissions_from_waste +
  greenhouse_gas_emissions_from_buildings +
  greenhouse_gas_emissions_from_industry +
  greenhouse_gas_emissions_from_manufacturing_and_construction +
  greenhouse_gas_emissions_from_transport +
  greenhouse_gas_emissions_from_electricity_and_heat +
  fugitive_emissions_of_greenhouse_gases_from_energy_production +
  greenhouse_gas_emissions_from_other_fuel_combustion +
  greenhouse_gas_emissions_from_bunker_fuels,
  data = gdp_sec_ghg)

summary(gdp_ghg_sec_reg)

kable(summary(gdp_ghg_sec_reg)$coefficients, caption = "GDP, Annual GHG and Sectorwise GHG Regression Coefficients")

#extracting values adjusted R, multiple R and f-statistics from the model 
adjusted_r_squared <- summary(gdp_ghg_sec_reg)$adj.r.squared
multiple_r <- summary(gdp_ghg_sec_reg)$r.squared
f_statistic <- summary(gdp_ghg_sec_reg)$fstatistic[1]
df1 <- summary(gdp_ghg_sec_reg)$fstatistic[2]
df2 <- summary(gdp_ghg_sec_reg)$fstatistic[3]
f_p_value <- pf(f_statistic, df1, df2, lower.tail = FALSE)

cat("Adjusted R-squared:", adjusted_r_squared, "\n")
cat("Multiple R:", multiple_r, "\n")
cat("F-statistic:", f_statistic, "\n")
cat("F-statistic p-value:", f_p_value, "\n")



```

```{r}
#Regression plot for Actual vs predicted GDP per capita 

gdp_sec_ghg$predicted_gdp_per_capita <- predict(gdp_ghg_sec_reg, newdata = gdp_sec_ghg)

# Plotting the actual values against the predicted values
ggplot(gdp_sec_ghg, aes(x = gdp_per_capita, y = predicted_gdp_per_capita)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "blue") +
  labs(title = "Actual vs. Predicted GDP per capita",
       x = "Actual GDP per capita",
       y = "Predicted GDP per capita")
```

#### Comparing GDP per capita, Annual CO2 emissions and sector

```{r}

#preparing the data 

#joining the datasets and rearranging the columns

gdp_sec_co2 <- left_join(gdp_pc, co2_sector, by = "Year") 
gdp_sec_co2

gdp_sec_co2 <- left_join(ann_co2, gdp_sec_co2, by = "Year")
gdp_sec_co2 <- clean_names(gdp_sec_co2)

```

```{r}
#renaming columns for convention

colnames(gdp_sec_co2) <- c("year", "annual_co2", "gdp_per_capita", "industry", "cement", "flaring", "gas", "oil", "coal")

gdp_sec_co2 <- gdp_sec_co2 %>%
  select(year, gdp_per_capita, everything())
```

```{r}

#checking if there are any missing values 

sum(is.na(gdp_sec_co2))

```

```{r}
#descriptive statistics

summary_gdp_sec_co2 <- data.frame(
  Variable = colnames(gdp_sec_co2)[-1],  # Exclude 'Year'
  N = nrow(gdp_sec_co2),
  Mean = sapply(gdp_sec_co2[-1], mean, na.rm = TRUE),
  Median = sapply(gdp_sec_co2[-1], median, na.rm = TRUE),
  SD = sapply(gdp_sec_co2[-1], sd, na.rm = TRUE),
  Min = sapply(gdp_sec_co2[-1], min, na.rm = TRUE),
  Max = sapply(gdp_sec_co2[-1], max, na.rm = TRUE)
)
print(summary_gdp_sec_co2)
```

#### Correlation Analysis for GDP per capita, Annual CO2 emissions and cumulative CO2 emissions from varios sources in NZ

```{r}

colnames(gdp_sec_co2)

gdp_sec_co2 <- clean_names(gdp_sec_co2)

 gdp_sec_co2_corr <- gdp_sec_co2 %>%
  select(gdp_per_capita, annual_co2, industry, cement, flaring, gas, oil, coal)

gdp_sec_co2_corr_matrix <- cor(gdp_sec_co2_corr, use = "complete.obs")
print(gdp_sec_co2_corr_matrix)

kable(gdp_sec_co2_corr_matrix, digits = 2, caption = "Correlation Matrix for GDP, Annnual CO2, and Cumulative CO2 emissions from multiple sectors in NZ")

```

Correlation Analysis

```{r}

correlation_gdp_ann_sec <- melt(gdp_sec_co2_corr_matrix)

ggplot(data = correlation_gdp_ann_sec, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "red", high = "green", mid = "yellow", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab",
                       name = "Correlation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1, size = 6, hjust = 1),
        axis.text.y = element_text(size = 6)) +
  labs(title = "Correlation Heatmap for GDP, Annual CO2 and Cumulative CO2 Emissions from various sources", x = "", y = "")


```

#### Regression Analysis

```{r}

colnames(gdp_sec_co2)

gdp_ann_sec_reg <- lm(gdp_per_capita ~ annual_co2 +
  industry + cement + flaring + gas + oil + coal,
  data = gdp_sec_co2)

summary(gdp_ann_sec_reg)

kable(summary(gdp_ann_sec_reg)$coefficients, caption = "GDP, Annual CO2 and Cumulative CO2 Emissions from various sources")

#extracting values adjusted R, multiple R and f-statistics from the model 
adjusted_r_squared <- summary(gdp_ann_sec_reg)$adj.r.squared
multiple_r <- summary(gdp_ann_sec_reg)$r.squared

cat("Adjusted R-squared:", adjusted_r_squared, "\n")
cat("Multiple R:", multiple_r, "\n")

```

```{r}

gdp_sec_co2$predicted_gdp_per_capita <- predict(gdp_ann_sec_reg, newdata = gdp_sec_co2)

# Plotting the actual values against the predicted values
ggplot(gdp_sec_co2, aes(x = gdp_per_capita, y = predicted_gdp_per_capita)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "blue") +
  labs(title = "Actual vs. Predicted GDP per capita",
       x = "Actual GDP per capita",
       y = "Predicted GDP per capita")
```

Forecasting variables for the next 5 years i.e. till 2027

#### Data mining for GDP per capita, Annual CO2 and Cumulative CO2 from different sources in NZ

```{r}

gdp_co2_dm <- read_csv("GDP-CO2-Sec.csv", show_col_types = FALSE)

pca_result <- prcomp(gdp_co2_dm, center = TRUE, scale. = TRUE)

summary(pca_result)

biplot(pca_result, main = "PCA of CO2 emissions and GDP Data")


```

#### autocorrelation and partial autocorrelation analysis

```{r}

#GHG per capita

ghg_pc_ac <- read.csv("GHG_per_capita.csv")

ghg_pc_ac <- ghg_pc_ac %>%
  filter(Entity == "New Zealand")
```

```{r}

ghg_per_capita_ts <- ts(ghg_pc_ac$Per.capita.greenhouse.gas.emissions.in.CO..equivalents, start=min(ghg_pc_ac$Year), end=max(ghg_pc_ac$Year), frequency=1)

autoplot(ghg_per_capita_ts) +
  ggtitle("GHG per Capita Emissions Over Time") +
  xlab("Year") +
  ylab("GHG per Capita")


#autocorrelation - AC

ggAcf(ghg_per_capita_ts) +
  ggtitle("Autocorrelation of GHG per Capita Emissions")

#partial autocorrelation - PAC
ggPacf(ghg_per_capita_ts) +
  ggtitle("Partial Autocorrelation of GHG per Capita Emissions")


```

```{r}
#GDP per capita 

gdp_pc_ac <- gdp_pc

```

```{r}

gdp_per_capita_ts <- ts(gdp_pc_ac$`GDP per capita`, start=min(gdp_pc_ac$Year), end=max(gdp_pc_ac$Year), frequency=1)

autoplot(gdp_per_capita_ts) +
  ggtitle("GDP per Capita Over Time") +
  xlab("Year") +
  ylab("GDP per Capita")


#autocorrelation - AC

ggAcf(gdp_per_capita_ts) +
  ggtitle("Autocorrelation of GDP per Capita")

#partial autocorrelation - PAC
ggPacf(gdp_per_capita_ts) +
  ggtitle("Partial Autocorrelation of GDP per Capita")



```

```{r}

#Annual GHG Emissions 

ghg_ac <- read_csv("Ann-GHG-NZ.csv", show_col_types = FALSE)

```

```{r}

ghg_ts <- ts(ghg_ac$`Annual greenhouse gas emissions in CO2 equivalents`, start=min(ghg_ac$Year), end=max(ghg_ac$Year), frequency=1)

autoplot(ghg_ts) +
  ggtitle("Annual GHG Over Time") +
  xlab("Year") +
  ylab("GHG")


#autocorrelation - AC
ggAcf(ghg_ts) +
  ggtitle("Autocorrelation of Annual GHG")

#partial autocorrelation - PAC
ggPacf(ghg_ts) +
  ggtitle("Partial Autocorrelation of Annual GHG")

```

```{r}
#Annual CO2 

co2_ac <- read_csv("Ann-CO2.csv", show_col_types = FALSE)

```

```{r}
co2_ts <- ts(co2_ac$`Annual CO2 emissions including land-use change`, start=min(co2_ac$Year), end=max(co2_ac$Year), frequency=1)

autoplot(co2_ts) +
  ggtitle("Annual CO2 Over Time") +
  xlab("Year") +
  ylab("CO2")


#autocorrelation - AC

ggAcf(co2_ts) +
  ggtitle("Autocorrelation of Annual CO2")

#partial autocorrelation - PAC
ggPacf(co2_ts) +
  ggtitle("Partial Autocorrelation of Annual CO2")


```

#### Predicting values for next 5 years using ARIMA

```{r}
#GDP per capita 

gdp_arima <- auto.arima(gdp_per_capita_ts)
summary(gdp_arima)

# Forecast for the next 5 years
gdp_forecast <- forecast(gdp_arima, h = 5)
plot(gdp_forecast)


```

```{r}
#Annual GHG


ghg_arima <- auto.arima(ghg_ts)
summary(ghg_arima)

# Forecast for the next 5 years
ghg_forecast <- forecast(ghg_arima, h = 5)
plot(ghg_forecast)


```

```{r}
#annual CO2

co2_arima <- auto.arima(co2_ts)
summary(co2_arima)

# Forecast for the next 5 years
co2_forecast <- forecast(co2_arima, h = 5)
plot(co2_forecast)


```

```{r}

#GHG Per capita

ghg_per_capita_arima <- auto.arima(ghg_per_capita_ts)
summary(ghg_per_capita_arima)

# Forecast for the next 5 years
ghg_per_capita_forecast <- forecast(ghg_per_capita_arima, h = 5)
plot(ghg_per_capita_forecast)



```

#### Evaluating model performance

```{r}
# Check residuals for GDP per Capita model
checkresiduals(gdp_arima)

# Repeat for other models
checkresiduals(ghg_arima)
checkresiduals(co2_arima)
checkresiduals(ghg_per_capita_arima)

```

```{r}
# Compare AIC values
AIC(gdp_arima, ghg_arima, co2_arima, ghg_per_capita_arima)

```

```{r}

thank_you_text <- data.frame(
  x = 1,
  y = 1,
  label = "Thank You!"
)

ggplot(thank_you_text, aes(x, y, label = label)) +
  geom_text(size = 20, color = "blue", fontface = "bold", family = "serif") +
  xlim(0.5, 1.5) +
  ylim(0.5, 1.5) +
  theme_void() +
  theme(
    plot.margin = margin(1, 1, 1, 1, "cm"),
    plot.background = element_rect(fill = "lightgray")
  )


```
